### 2018-03-20 ComputationalLinearAlgebra/Homework/Homework1.tex
what da hell

\documentclass{article}
\input{../../definitions.tex}

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}

\title{Computational Linear Algebra\\ Homework 1}
\author{Benji Altman}

\begin{document}
	\maketitle
	\tableofcontents

	\section{Question 1}
	See attached R code.

	\section{Question 2}
	\subsection{}
	See attached R code.

	\subsection{}
	See attached R code.
	\subsection{}
	Let $f(x) = \cos(x)^2 + \sin(x) - 1$, then we find that $f'(x) = - 2\sin(x)\cos(x) + \cos(x)$. Then the tangent line at $x_0$ for any $x_0\in\mathbb R$ to $f$ is 
	\begin{align*}
	\ell(x) &= f'(x_0)(x-x_0) + f(x_0)\text{.}
	\end{align*}
	We now solve for $x$ when $\ell(x) = 0$.
	\begin{align*}
	\ell(x) &= 0 \\
	f(x_0)(x-x_0)+f'(x_0) &= 0\\
	f(x_0)(x-x_0) &= -f'(x_0) \\
	x-x_0 &= -\frac{f'(x_0)}{f(x_0)}\\
	x  &= x_0-\frac{f'(x_0)}{f(x_0)}\\
	&=x_0+\frac{2\sin(x_0)\cos(x_0)-\cos(x_0)}{\cos(x_0)^2 + \sin(x_0)-1}
	\end{align*}

	This means that in newton's method for $f$,
	\[
	x_{i+1} = x_i+\frac{2\sin(x_i)\cos(x_i)-\cos(x_i)}{\cos(x_i)^2 + \sin(x_i)-1}\text{.}
	\]
	\subsection{}
	See attached R code.
	\subsection{}
	See attached R code.
	\section{Question 3}
	We have the matrix
	\[
	A = 
	\begin{bmatrix}
	1&2&3&x\\
	4&5&x&6\\
	7&x&8&9\\
	x&10&11&12
	\end{bmatrix}
	\]
	and we wish to solve for $x$ when $\det A = 1000$. We will then let $f(x) = \det A - 1000$ and we may solve for some roots.

	The rest of this problem will be doe in R; see attached R code.

	\section{Question 4}
	See attached R code.
\end{document}