### 2018-03-20 ComputationalLinearAlgebra/Homework/Homework3.Rmd

---
title: "Homework 3"
author: "Benji Altman"
date: "February 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### Part a

So first notice that $L_k = I - E_k$ for some $n\times n$ matrix $E_k$. Notice that $$E_k = 
\begin{bmatrix}
0 \\
 & \ddots \\ 
 & & 0 \\
 & & \ell_{k+1,k} & 0 \\
 & & \ddots& & \ddots \\
 & & \ell_{n,k} & & & 0
\end{bmatrix}
$$

Notice now that $\ell_ke_k^T = E_k$ as $f(\vec v) = e_k^T \vec v$ will take a column vector $\vec v$ and put it into the $k^{\text{th}}$ column of a matrix. Additionally we would like to notice that ${E_k}^2 = 0$ as $E_k$ has zeros on the diagonal and all indicies above it, thus for any index $i,j$ in ${E_k}^2$ the dot product that gives it would contain a zero in either the first or second vector in every index.

Now let $A$ be the matrix that we would like to show is $T_k^{-1}$
\begin{align*}
L_k &= I-E_k \\
A &= I+E_k \\
L_k A &= (I-E_k)(I+E_k) \\
&= I^2-(E_k)^2 \\
&= I
\end{align*}
And now by definition $A$ is the inverse of $T_k$.

### Part b

Now notice that 
\begin{align*}
\prod_{k=1}^nT_k^{-1} &= \prod_{k=1}^n (I+E_k) \\
&= (I+E_1)\cdot(I+E_2)\cdots(I+E_n) \\
&= (I^2+E_1 + E_1 + E_2 + E_1\cdot E_2)(I+E_3)\cdots(I+E_n) \\
&= (I + E_1 + E_2)(I+E_3)\cdots(I+E_n) \\
&= \ldots \\
&= I + E_1 + E_2 + \ldots + E_n
\end{align*}

Again notice that $E_aE_b = 0_{n\times n}$ as $E_a$ and $E_b$ have zeros on the diagonal and everywhere above it.

## Question 2

### Part a

$$A = 
\begin{bmatrix}
4&-1&-1&0&0&0&0&0 \\
-1&4&0&-1&0&0&0&0 \\
-1&0&4&-1&-1&0&0&0\\
0&-1&-1&4&0&-1&0&0\\
0&0&-1&0&4&-1&-1&0\\
0&0&0&-1&-1&4&0&-1\\
0&0&0&0&-1&0&4&-1\\
0&0&0&0&0&-1&-1&4\\
\end{bmatrix}
$$
 and

$$ 
b = 
 \begin{bmatrix}
 5 \\ 15 \\ 0 \\ 10 \\ 0 \\ 10 \\ 20 \\ 30
 \end{bmatrix}
$$

### Part b
```{r}
#Takes in matrix, A, and does a L-U decomposition
#Complete, tested
LUDecompose = function(A)
{
  if(nrow(A) != ncol(A))
    return(NULL)
  out = new.env()
  out$U = A
  out$L = diag(nrow(A))

  #now lets start getting to it
  i = 1
  while(i < nrow(A))
  {
    #make ith column 0
    row = i + 1
    while(row <= nrow(A))
    {
      #set row
      op = out$U[row, i] / out$U[i, i]
      out$L[row, i] = op
      out$U[row,] = out$U[row,] - out$U[i,] * op
      row = row + 1
    }
    i = i + 1
  }
  return(out)
}
#Now to type in A
A = matrix(
  c(4,-1,-1,0,0,0,0,0,
    -1,4,0,-1,0,0,0,0,
    -1,0,4,-1,-1,0,0,0,
    0,-1,-1,4,0,-1,0,0,
    0,0,-1,0,4,-1,-1,0,
    0,0,0,-1,-1,4,0,-1,
    0,0,0,0,-1,0,4,-1,
    0,0,0,0,0,-1,-1,4),
  nrow = 8, ncol = 8, byrow = TRUE)

factorization = LUDecompose(A)
print(factorization$U)
print(factorization$L)
```

### Part c

```{r, echo=FALSE}
mySolve=function(L,U,b,tol=1e-10){

  n=nrow(L)

  # First solve Ly=b 
  y = rep(0,n)        # pre-allocate a vector y with 0s in it.
  if(abs(L[1,1])<tol) stop('There is a zero on the diagonal of L')
  y[1] = b[1]/L[1,1]  # Fill in the 1st value of y
  for (j in 2:n ) {
    if(abs(L[j,j])<tol) stop('There is a zero on the diagonal of L')
    y[j] = (b[j] - L[j,1:(j-1)]%*%y[1:(j-1)])/L[j,j]
  }

  # Then solve Ux=y 
  x = rep(0,n)        # pre-allocate a vector x with 0s in it.
  if(abs(U[n,n])<tol) stop('There is a zero on the diagonal of U')
  x[n] = y[n]/U[n,n]  # Fill in the nth value of x
  for (j in (n-1):1 ) {
    if(abs(U[j,j])<tol) stop('There is a zero on the diagonal of U')
    x[j] = (y[j] - U[j,(j+1):n]%*%x[(j+1):n])/U[j,j]
  }
  return(x)
}
```

```{r}
mySolve(factorization$L, factorization$U, c(5,15,0,10,0,10,20,30))
```

### Part d

```{r}
mySolve(factorization$L, factorization$U, c(5,15,0,10,0,10,30,40))
```

### Part e

```{r}
solve(A)
```

## Question 6

Take a matrix $A$, one may define $x_i$ such that $x_i$ maximizes the $i^{\text{th}}$ value in $Ax_i$. One does this by letting ${x_i}_j = \begin{cases} 1 & A_{i,j}\ge0\\-1&A_{i,j}<0\end{cases}$. Now we may maximize $||Ax_i||_\infty$, by letting $i$ be the row in $A$'s sum of absolute values is greatest. This means that $|| A||_\infty$ is the maximum sum of absolute values of rows

## Question 7

let
$$
A=\begin{bmatrix}
-1 & 0 & 1 \\
3 & 1 & 1 \\
2 & 0 & 1 
\end{bmatrix} 
$$
Now we will do this in steps

$$
P_1 = \begin{bmatrix}
0&1&0\\
1&0&0\\
0&0&1
\end{bmatrix}
$$
$$
A_1 = \begin{bmatrix}
3&1&1\\
-\frac13&\frac13&\frac43\\
\frac23&-\frac23&\frac13
\end{bmatrix}
$$
$$
P_2 = \begin{bmatrix}
0&1&0\\
0&0&1\\
1&0&0
\end{bmatrix}
$$
$$
A_2 = \begin{bmatrix}
3&1&1\\
\frac23&-\frac23&\frac13\\
-\frac13&-\frac12&\frac32
\end{bmatrix}
$$
Thus we have 
$$
\begin{bmatrix}
0&1&0\\0&0&1\\1&0&0
\end{bmatrix}
\begin{bmatrix}
-1 & 0 & 1 \\
3 & 1 & 1 \\
2 & 0 & 1 
\end{bmatrix}
=
\begin{bmatrix}
1&0&0\\\frac23&1&0\\-\frac13&-\frac12&1
\end{bmatrix}
\begin{bmatrix}
3&1&1\\0&-\frac23&\frac13\\0&0&\frac32
\end{bmatrix}
$$