### 2018-03-20 MathPapers/Seminar/DFAs.tex

\documentclass{article}
\input{../../definitions}
\title{Deterministic Finite Automata}
\begin{document}
	\maketitle

	\question{1.} Make a DFA for the language of all binary integers.

	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{2.} Make a DFA for the language of all binary numbers with finite representation (only rational numbers that do not need repeating representation).
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{3.} Make a DFA for numbers written in scientific notation.
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{4.} Make a DFA for the language on $\set{a,b}$ with no two $b$'s in a row.
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{5.} Make a DFA for the language on $\set{a,b}$ with no $b$ by itself.
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{6.} Make a DFA for the language $\set{\lambda, a, aa, aab, aaba, aabaa, aabaab, \ldots}$
\end{document}