### 2018-03-20 Topology/Paper/glossary.tex

\usepackage[toc,xindy]{glossaries}

\makenoidxglossaries

\newglossaryentry{set}
{
	name={set},
	description={A collection of objects}
}
\newglossaryentry{vertex}
{
	name={vertex},
	description={A point or node in a graph},
	plural={verticies}
}
\newglossaryentry{edge}
{
	name={edge},
	description={A connection or line between verticies in a graph}
}
\newglossaryentry{connected}
{
	name={connected},
	description={A graph where one may start at any point and could follow edges and eventually get to any other point}
}
\newglossaryentry{multigraph}
{
	name={multigraph},
	description={Like a graph, but multiple edges may connect the same vertex pair, and an edge may connect a vertex to itself}
}
\newglossaryentry{complete}
{
	name={complete graph},
	description={A graph with all possible edges included, the notation $K_n$ is used to denote the complete with $n$ verticies}
}
\newglossaryentry{face}
{
	name={face},
	description={An section of the space we are embedding our graph in that is separated from the rest of the space by edges}
}
\newglossaryentry{contraction}
{
	name={contraction},
	description={A graph operation where one removes an edge by fusing two verticies together}
}
\newglossaryentry{subgraph}
{
	name={subgraph},
	description={$A$ is a subgraph of some graph $\mathcal G$ iff you could add verticies and edges to $A$ and somehow get $\mathcal G$}
}
\newglossaryentry{graph}
{
	name={graph},
	description={A set of verticies and edges}
}