### 2018-03-27 Algebra/Homework/Homework.tex
upd

\documentclass{article}

\input{../../definitions.tex}
\input{../algebra.tex}

\title{Algebra Homework}

\begin{document}
\maketitle
\tableofcontents

\section{Chapter 1}
\subsection{Section 1}
\subsubsection{Question 1}

Choose $a,b\in S$. We find $$a = a * b = b * a = b$$, and thus all elements in $S$ must be the same element, so there is most one element of $S$.

\subsubsection{Question 2}
Let us choose $a,b,c \in S$.

\question{(a)} We have $$a*b = a - b = -(b-a) = -(b*a)$$, thus iff $0 = a*b = a-b$ we have $a*b = b*a$ as $0=-0$, however for any other value of $a*b$, $a*b\not=b*a$. We also may notice that iff $a=b$, then $a*b=a-b=0$. Thus for all $a\not= b$, $a*b \not=b*a$.

\question{(b)} We have
\begin{align*}
a*(b*c) &= a-(b-c) \\
&= a+(c-b) \\
&= a+c-b \\
&= a-b+c \\
&= a-b-(-c) \\
&= (a-b)-(-c) \\
&= (a*b)*-c
\end{align*} so $a*(b*c) = (a*b)*c$ iff $c=-c$ which is only true if $c = 0$.

\question{(c)} We have $a*0 = a-0 = a$.

\question{(d)} We have $a*a = a-a = 0$.

\subsection{Section 2}

\subsubsection{Question 8}

Let $x \in (A\setminus B) \union (B\setminus A)$ then either $x \in A \setminus B$ or $x \in B\setminus A$. If $x \in A\setminus B$ then we get that $x \in A$ and $x \not\in B$, thus $x \in A\union B$ and $x \not\in A \inter B$, which would mean $x \in (A\union B)\setminus(A\inter B)$. If $x\in B\setminus A$ then we get that $x\in B$ and $x\not\in A$, thus $x\in A \union B$ and $x\not\in A\inter B$, which would mean $x \in (A\union B)\setminus(A\inter B)$. It has now been demonstrated that $(A\setminus B) \union (B\setminus A) \subset (A \union B) \setminus (B\inter A)$.

Now let $x\in(A\union B)\setminus(A\inter B)$. We have that $x \in A\union B$ and $x \not\in A\inter B$. It follows that either $x \in A$ or $x\in B$, however, $x$ is not in both $A$ and $B$. This may be written as: $x \in A$ and $x \not\in B$, or $x\in B$ and $x\not\in A$. This then translates to $x \in A\setminus B$ or $x\in B\setminus A$, therefore, $x\in (A\setminus B) \union (B\setminus A)$. It has now been demonstrated that $(A\union B) \setminus (B\inter A) \subset (A\setminus B)\union (B\setminus A)$.

Now it has been shown that both sets are subsets of each-other, thus $(A\setminus B)\union(B\setminus A) = (A\union B) \setminus (A\inter B)$.

This may be displayed pictorially as follows:

%TODO figure out pictures

\def\secondcircle{(210:0.95cm) circle (1.5cm)}
\def\thirdcircle{(330:0.95cm) circle (1.5cm)}
\begin{tikzpicture}
	\begin{scope}
		\clip \secondcircle;
		\fill[cyan] \thirdcircle;
	\end{scope}
	\draw \secondcircle node [text=black,below left] {$A$};
	\draw \thirdcircle node [text=black,below right] {$B$};
\end{tikzpicture}

\subsubsection{Question 9}

Let $x \in A \inter (B \union C)$, thus $x \in A$ and $x\in B\union C$. We then have that $x \in B$ or $x \in C$. Now as we already know that $x\in A$ then we get that either $x \in B \inter A$ or $x \in C \inter A$ and therefore $x \in (A\inter B) \union (A \inter C)$. Thus it has been shown that $A \inter (B\union C) \subset (A\inter B)\union (A\inter C)$.

Let $x \in (A\inter B) \union (A\inter C)$, thus $x \in (A\inter B)$ or $x\in (A\inter C)$. We then get that either $x \in A$ and $x \in B$ or that $x\in A$ and $x\in C$, either way $x\in A$, thus we may write that $x \in A$ and either $x\in B$ or $x\in C$. This would be the same as $x\in A$ and $x\in B\union C$, which then translates to $x \in A \inter (B\union C)$. Thus it has been shown that $(A\inter B)\union (A\inter C) \subset A\inter (B\union C)$.

We have now shown that both sets are subsets of each-other, thus $A\inter(B\union C) = (A\inter B)\union (A\inter C)$.

\subsubsection{Question 10}

Let $x\in A\union (B\inter C)$, assume then for the sake of contradiction that $x\not\in(A\union B)\inter (A\union C)$. Because $x \in A\union (B\inter C)$ we have that $x \in A$ or $x\in B\inter C$. Because $x\not\in (A\union B)\inter (A\union C)$ we have that $x\not\in A\union B$ or $x\not\in A\union C$. We then get that either $x \not\in A$ and $x\not\in B$ or $x\not\in A$ and $x\not\in C$, either way $x\not\in A$, so we have $x\in B\inter C$. We know that $x\not\in B$ or $x\not\in C$, however we also have that $x\in B$ and $x\in C$ due to $x\in B\inter C$, thus we have a contradiction. Thus $A\union(B\inter C) \subset (A\union B)\inter(A \union C)$.

Let $x\in (A \union B) \inter (A\union C)$ and assume for the sake of contradiction that $x\not\in A\union(B \inter C)$. We then get that $x\not\in A$ and $x\not\in B\inter C$. We also have that $x\in A \union B$ and $x\in A \union C$, so if $x\not\in A$ then we get $x\in B$ and $x\in C$. This is then translated to $x\in B\inter C$ which is a direct contradiction with $x\not\in B\inter C$ and again we have a contradiction. Thus $(A\union B)\inter(A\union C)\subset A\union (B\inter C)$.

We have now shown that both sets are subsets of each other, thus $A\inter(B\union C) = (A\union B)\inter(A\union C)$.

\subsubsection{Question 12}

\question{(a)}
\begin{align*}
	\setcomp{(A\union B)} &= \setbuilder{x\in S}{x\not\in A\union B} \\
	&= \setbuilder{x\in S}{x\not\in A \And x\not\in B} \\
	&= \setbuilder{x\in S}{x \in \setcomp A \And x\in \setcomp B} \\
	&= \setcomp A \inter \setcomp B
\end{align*}

\question{(b)}
\begin{align*}
	\setcomp{(A\inter B)} &= \setbuilder{x\in S}{x\not\in A\inter B} \\
	&= \setbuilder{x\in S}{x\not\in A \Or x\not\in B} \\
	&= \setbuilder{x\in S}{x \in \setcomp A \Or x\in \setcomp B} \\
	&= \setcomp A \union \setcomp B
\end{align*}

\subsubsection{Question 13}
\question{(a)}
\begin{align*}
	A + B &= (A\setminus B) \union (B\setminus A) \\
	&= (B\setminus A) \union (A\setminus B) \\
	&= B + A
\end{align*}

\question{(b)}
First notice that for any set $X$, $X\setminus \emptyset = A$ and that $\emptyset \setminus X = \emptyset$.
\begin{align*}
	A + \emptyset &= (A \setminus \emptyset) \union (\emptyset \setminus A)\\
	&= A \union \emptyset \\
	&= A \\
\end{align*}

\question{(c)}
\begin{align*}
	A\cdot A &= A\inter A\\
	&= A
\end{align*}

\question{(d)}
\begin{align*}
	A + A &= (A\setminus A)\union (A\setminus A) \\
	&= \emptyset \union \emptyset \\
	&= \emptyset
\end{align*}

\question{(e)} To simplify this question let me introduce the logical operation, $a \lxor b$ which is defined as either $a$ or $b$ but not both, and we will show that $a \lxor (b\lxor c) = (a\lxor b)\lxor c$ using truth tables.

\begin{center}
	\begin{tabular}{c|c|c|c|c|c|c}
		$a$&$b$&$c$&$a\lxor b$&$b\lxor c$&$a\lxor(b\lxor c)$&$(a\lxor b)\lxor c$ \\
		\false&\false&\false&\false&\false&\false&\false\\
		\false&\false&\true&\false&\true&\true&\true\\
		\false&\true&\false&\true&\true&\true&\true\\
		\false&\true&\true&\true&\false&\false&\false\\
		\true&\false&\false&\true&\false&\true&\true\\
		\true&\false&\true&\true&\true&\false&\false\\
		\true&\true&\false&\false&\true&\false&\false\\
		\true&\true&\true&\false&\false&\true&\true
	\end{tabular}
\end{center}

Now we wish to show that $A + B = \setbuilder{x\in S}{x \in A \lxor x \in B}$. To do this we will first show that $a\lxor b = (a\land\lnot b) \lor (b\land\lnot a)$, where $\lnot$ is a logical not, $\land$ is a logical and, and $\lor$ is a logical or. We again show this by the following truth table:
\begin{center}
	\begin{tabular}{c|c|c|c|c|c|c|c}
		$a$&$b$&$\lnot b$&$a\land\lnot b$&$\lnot a$&$b\land\lnot a$&$(a\land\lnot b)\lor(b\land\lnot a)$&$a\lxor b$\\
		\false&\false&\true&\false&\true&\false&\false&\false\\
		\false&\true&\false&\false&\true&\true&\true&\true\\
		\true&\false&\true&\true&\false&\false&\true&\true\\
		\true&\true&\false&\false&\false&\false&\false&\false\\
	\end{tabular}
\end{center}
Now we find
\begin{align*}
A + B &= \setbuilder{x\in S}{x\in A+B} \\
&= \setbuilder{x\in S}{x \in (A\setminus B)\union(B\setminus A)}\\
&= \setbuilder{x\in S}{x\in (A\setminus B) \lor x\in(B\setminus A)}\\
&=\setbuilder{x\in S}{(x\in A \land x\not\in B)\lor(x\in B\land x\not\in A)}\\
&=\setbuilder{x\in S}{x\in A\lxor x\in B}\\
\end{align*}
so we then have
\begin{align*}
A+(B+C)&=\setbuilder{x\in S}{x\in A\lxor x\in B+C} \\
&=\setbuilder{x\in S}{x\in A\lxor(x\in B\lxor x\in C)}\\
&=\setbuilder{x\in S}{(x\in A \lxor x\in B)\lxor x\in C}\\
&=\setbuilder{x\in S}{x \in A + B\lxor x \in C}\\
&=(A+B)+C
\end{align*}

\question{(f)}
Suppose $B\not= C$. Because $B\not= C$ there exists some $x\in S$ such that either $x\in B$ and $x\not\in C$ or $x\in C$ and $x\not\in B$, we will assume without loss of generality that $x\in B$ and $x\not\in C$. Now if $x\in A$ then we would find $x\not\in A+B$ and $x\in A+C$. If $x\not\in A$ we would find that $x\in A+B$ and $x\not\in A+C$. We now have shown that $B\not= C\implies A+B\not=A+C$, thus by contrapositive we have $A+B=A+C \implies B=C$.

\question{(g)}
First we will want to show logical equivalence between the statement $a\land(b\lxor c)$ and $(a\land b)\lxor(a\land c)$.
\begin{center}
	\begin{tabular}{c|c|c|c|c|c|c|c}
		$a$&$b$&$c$&$b\lxor c$&$a\land b$&$a\land c$&$a\land(b\lxor c)$&$(a\land b)\lxor(a\land c)$\\
		\false&\false&\false&\false&\false&\false&\false&\false\\
		\false&\false&\true&\true&\false&\false&\false&\false\\
		\false&\true&\false&\true&\false&\false&\false&\false\\
		\false&\true&\true&\false&\false&\false&\false&\false\\
		\true&\false&\false&\false&\false&\false&\false&\false\\
		\true&\false&\true&\true&\false&\true&\true&\true\\
		\true&\true&\false&\true&\true&\false&\true&\true\\
		\true&\true&\true&\false&\true&\true&\false&\false
	\end{tabular}
\end{center}
now we may show
\begin{align*}
A\cdot(B+C) &= A\inter(B+C)\\
&=\setbuilder{x\in S}{x \in A\inter(B+C)}\\
&= \setbuilder{x\in S}{x\in A\land x\in(B+C)} \\
&= \setbuilder{x\in S}{x\in A\land (x\in B \lxor x\in C)}\\
&=\setbuilder{x\in S}{(x\in A\land x\in B)\lxor(x\in A\land x\in C)}\\
&=\setbuilder{x\in S}{x\in A\inter B\lxor x\in A\inter C}\\
&=\setbuilder{x\in S}{x\in (A\inter B)+(A\inter C)}\\
&=(A\inter B)+(A\inter C)\\
&=(A\cdot B)+(A\cdot C)
\end{align*}

\subsubsection{Question 14}
First notice that if $A$ and $B$ are disjoint then $m(A\union B) = m(A) + m(B)$. So now we get the three disjoint sets $A\setminus B$, $A\inter B$, and $B\setminus A$, notice that $A =( A\setminus B) \union (A\inter B)$, that $B =(B\setminus A)\union(A\inter B)$, and $A\union B = (A\setminus B)\union (A\inter B)\union (B\setminus A)$. Now we get $m(A) = m(A\setminus B) + m(A\inter B)$, $m(B) = m(B\setminus A) + m(A\inter B)$, and $m(A\union B) = m(A\setminus B)+(A\inter B)+m(B\setminus A)$. We then get \begin{align*}
m(A) + m(B) &= m(A\setminus B)+m(A \inter B)+m(B\setminus A)+m(A\inter B)\\
&= m(A\union B) + m(A\inter B)\\
m(A) + m(B) - m(A \inter B) &= m(A \union B)
\end{align*}

\subsubsection{Question 22}
\question{(a)} To construct a subset of any set we go through each element and choose to include it or not to, this gives us two possibilities per element. For a set of size $n$ then there are $n$ independent choices to be made in constructing a subset, thus $2^n$ subsets.

\question{(b)} There are exactly $\binom nm = \frac{n!}{m!(n-m)!}$ subsets of a set with $n$ elements that have $m$ elements.

\begin{proof}
	Let us start by defining $\binom nm$ as the number of ways to choose a subset with $m$ elements from a set with $n$ elements. Now we must recognize that $k!$ is the number of ways to order a set with $k$ elements. Then we get that $\binom nmm!(n-m)! = n!$ as we may order our set with $n$ elements by choosing the first $m$ elements in our order ($\binom nm$ possible ways), then ordering those elements ($m!$ ways), and finally ordering the rest of the elements ($(m-n!)$ ways). This gives us $\binom nmm!(n-m)!=n!$ and from there we divide and get $\binom nm=\frac{n!}{m!(n-m)!}$.
\end{proof}

\subsection{Section 3}
\subsubsection{Question 7}
Let $g:S\to T$, $h:S\to T$ and $f:T\to U$ be functions such that $f$ is 1-1 and $f\composed g = f\composed h$. Assume for the sake of contradiction that $g\not=h$, then there exists some $s \in S$ such that $g(s) \not= h(s)$. We know that $f\composed g(s) = f\composed h(s)$, thus $f(g(s)) = f(h(s))$ so $g(s) = h(s)$ by $f$ being 1-1. Thus we have a contradiction and we know that $g=h$.

\subsubsection{Question 8}
\question{(a)} Yes, as all integers are either even or odd and none are both even and odd.

\question{(b)} Let us break this into cases:
\begin{itemize}
	\item If $s_1$ and $s_2$ are even, then there exists $k_1\in\mathbb Z$ and $k_2\in\mathbb Z$ such that $2k_1 = s_1$ and $2k_2 = s_1$. Thus $s_1 + s_2 = 2k_1 + 2k_2 = 2(k_1 + k_2)$, thus $f(s_1 + s_2) = 1$. We also find that $f(s_1) \cdot f(s_2) = 1 \cdot 1 = 1$.
	\item If $s_1$ is even and $s_2$ is odd, then there exists $k_1\in\mathbb Z$ and $k_2\in\mathbb Z$ such that $s_1 = 2k_1$ and $s_2 = 2k_2 + 1$. Thus $s_1 + s_2 = 2k_1 + 2k_2 + 1 = 2(k_1 + k_2) + 1$ so $f(s_1 + s_2) = -1$. We also find that $f(s_1) f(s_2) = 1\cdot -1 = -1$.
	\item If $s_1$ is odd and $s_2$ is even we may write that $f(s_1 + s_2) = f(s_2 + s_1)$ and that $f(s_1)f(s_2) = f(s_2)f(s_1)$ because both addition and multiplication are commutative. Now we see that we have reproduced our previous case and thus in this case the equality holds.
	\item If $s_1$ and $s_2$ are odd, then there exists $k_1\in\mathbb Z$ and $k_2\in\mathbb Z$ such that $2k_1 + 1 = s_1$ and $2k_2 + 1 = s_2$, thus $s_1 + s_2 = 2k_1 + 1 + 2k_2 + 1 = 2(k_1 + k_2 + 1)$ so $f(s_1 + s_2) = 1$. We also find that $f(s_1)f(s_2) = -1\cdot -1 = 1$.
\end{itemize}
Thus for all possible integers $s_1$ and $s_2$, we have $f(s_1 + s_2) = f(s_1)f(s_2)$.

This tells us that even integers are closed under addition. that odd integers added together always are even, and finally that an odd added to an even is odd.

\question{(c)} No, as $f(1\cdot 2) = f(2) = 1$ and $f(1)f(2) = -1 \cdot 1 = -1$.

\subsubsection{Question 12}

\question{(a)} No $f$ is not a function as $2/3 = 4/6$ and $f(2/3) = 2^23^3 \not= 2^43^6 = f(4/6)$.

\question{(b)} We may define $f(m/n) = 2^m3^n$ iff $m$ and $n$ are coprime.

\subsubsection{Question 19}

Let $f(x) = x^2+ax+b$, thus $f'(x) = 2x+a$. $f'(x)$ is linear so there exists only one $x\in\mathbb R$ for which $f'(x) = 0$, and thus this $x$ is a global extrema for $f$, so $f$ can not be surjective. Now consider $x_1 = -\frac a2-1$ and $x_2=-\frac a2+1$, thus
\begin{align*}
f(x_1) &= \left(-\frac a2-1\right)^2+a\left(-\frac a2-1\right)+b \\
&= \frac{a^2}4+2\frac a2+1-\frac{a^2}2-a+b\\
&=\frac{a^2}4+1+b\\
f(x_2) &= \left(-\frac a2+1\right)^2+a\left(-\frac a2+1\right)+b \\
&= \frac{a^2}4-2\frac a2+1-\frac{a^2}2+a+b\\
&= \frac{a^2}4+1+b
\end{align*}

so $f$ must be 1-1.

\subsubsection{Question 23}

\def\RB{}
\def\LB{}

\marginnote{Ugly proof:}
First let us show that there exists some bijection from $\mathbb N$ to $\ZNonNegative^2$. Consider the 1 norm on $\ZNonNegative^2$, defined as $\norm{(a,b)}_1 = a + b$. Then we may partition $\ZNonNegative^2$ into subsets $P_n = \setbuilder{x\in\ZNonNegative^2}{\norm x_1 = n}$, for any $n\in \ZNonNegative$. Notice that for $(a,b)\in P_n$ then $a \le n$ and $b \le n$, thus forcing $P_n$ to be finite. Now we can construct a function mapping from $\mathbb N$ to $\ZNonNegative^2$ by giving each element of $P_1$ a number from $1$ to $\card{P_0}$ (inclusive), then the next $\card{P_1}$ will be given to elements of $P_1$ and so on infinitely. Notice that by construction $x\not= y \implies f(x) \not= f(y)$, so we get this being 1-1, additionally for any $(a,b) \in \ZNonNegative^2$, $(a,b)\in P_{a+b}$ and thus receives a number greater than $\sumfrom{n=0}{a+b-1}{\card{P_n}}$ and less than or equal to $\sumfrom{n=0}{a+b}{\card{P_n}}$. This means that we can label each element of $\ZNonNegative$ with a single natural number and thus have a bijection.

Now we can also construct a trivial bijection, $h:\ZNonNegative^2\to S$ as $h(a,b) = 2^a3^b$. Now we may compose the bijections to get a 1-1 correspondence $\mathbb N = S$ onto $T$.

\bigskip

\marginnote{Nice proof:}
First notice that $T \subset S$ so there exists the trivial injective function from $T$ to $S$. Second notice that $f:S\to T$ defined as $f(s)=2^s$ is both well defined as injective. By the  \href{https://en.wikipedia.org/wiki/Schr\%C3\%B6der\%E2\%80\%93Bernstein_theorem}{Schr\"oder-Bernstein theorem} there must be some bijection from $S$ to $T$.

\subsubsection{Question 28}

Let $S$ be a finite set, with $f:S\to S$. Now let $f(x) = f(y)$, for some $x\not=y$, then there remain $\card S - 2$ elements in $S\setminus\set{x,y}$ and $\card S - 1$ elements in $S\setminus\set{f(x)}$. This means that for any definition of $f$ on $S\setminus\set{x,y}$ it can not possibly be onto $S\setminus\set{f(x)}$. We have now shown $f$ not being 1-1 implies $f$ not being onto, by contrapositive $f$ being onto implies $f$ is 1-1.

\subsubsection{Question 29}

Let $S$ be a finite set, with $f:S\to S$ injective. Now as $f$ is 1-1 each $s\in S$ has a unique $f(s)\in S$, so $f(S)$ must have exactly $\card S$ unique elements, thus $f(S) \subset S$ with exactly $\card S$ elements.\footnote{$f(A)$ is defined as $\setbuilder{y\in\range{f}}{\Exists{x\in \domain{f}}{f(x) = y}}$ when $A \subset \domain{f}$ and $A\not\in\domain{f}$.} Because $S$ is finite, this implies $f(S) = S$.

\subsection{Section 4}

\subsubsection{Question 5}

\question{(a)} First identity:
\begin{align*}
f^2g^2&=ffgg\\
&=f(fg)g\\
&=f(gf)f\\
&=(fg)^2
\end{align*}

\question{(b)} Second Identity: Let $i$ be the identity function.
\begin{align*}
\inv f\inv ggf &= i\\
\inv f\inv ggf\inv{(gf)} &= i\inv{(gf)}\\
=\inv f\inv g i&=i\inv{(fg)}\\
=\inv f\inv g&=\inv{(fg)}
\end{align*}

\subsubsection{Question 9}

\question{(a)} 
\begin{align*}
f^2&:x_1\to x_3, x_2 \to x_4, x_3 \to x_1, x_4 \to x_2\\
f^3&:x_1\to x_4, x_2 \to x_1, x_3 \to x_2, x_4 \to x_1\\
f^4&:x_1\to x_1, x_2 \to x_2, x_3 \to x_3, x_4 \to x_4
\end{align*}

\question{(b)}
\begin{align*}
g^2&:x_1\to x_1, x_2 \to x_2, x_3 \to x_3, x_4 \to x_4\\
g^3&:x_1\to x_2, x_2 \to x_1, x_3 \to x_3, x_4 \to x_4
\end{align*}

\question{(c)}
$$fg:x_1\to x_3, x_2\to x_2, x_3\to x_4, x_4\to x_1$$

\question{(d)}
$$gf:x_1\to x_1, x_2 \to x_3, x_3 \to x_4, x_4 \to x_2$$

\question{(e)}
\begin{align*}
(fg)^3&:x_1\to x_1, x_2 \to x_2, x_3 \to x_3, x_4 \to x_4\\
(gf)^3&:x_1\to x_1, x_2 \to x_2, x_3 \to x_3, x_4 \to x_4
\end{align*}

\question{(f)} No, $fg(x_1) \not= gf(x_1)$ as can be seen above, thus $fg\not=gf$.

\subsubsection{Question 10}
Consider the cycle structure of a permutation $f$. It is obvious that $f^k = i$ if $k$ is the greatest common divisor among all the cycle lengths in $f$. Now for any $f\in S_3$, cycles must be of length one, two, or three. Therefore, as $6 = \gcd(1,2,3)$ for any $f\in S_3$, $f^6=i$.

\subsubsection{Question 14}

Let $F$ be the mapping from $S_m\to S_n$ such that $F(f)$ is defined to be the same as $f$ where $f$ is defined, and acts as the identity elsewhere. Now $F$ is trivially 1-1, so let us show that it satisfies $F(fg) = F(f)F(g)$ for all $f,g \in S_m$. To start let us choose $x$ in the domain of $g$, then $F(g)$ takes $x \to g(x)$ and $F(f)$ takes $g(x) \to fg(x)$, which is obviously the same as what $F(fg)$ does. If $x$ is not in the domain of $g$ then $F(g)$ takes $x\to x$ and $F(f)$ takes $x \to x$ as does $F(fg)$, we can thus conclude that $F(fg) = F(f)F(g)$.

\subsubsection{Question 21}

Let $g_j$ swap $x_1$ and $x_{j+1}$. Now when $n = 1$ this is trivially true as we have $f = i$ which satisfies the definition of $f$. Let us now try and do an induction on this statement. Assume that $g_1g_2g_3\cdots g_{n-1} = f$ when $n$ is some specific fixed constant. Then it follows that for $f'\in S_{n+1}$ where $f'$ is defined just as $f$ was, that is $f': x_1\to x_2, x_2 \to x_3, \ldots, x_n \to x_{n+1}, x_{n+1} \to x_1$, then consider $g_1g_2g_3 \ldots g_n = fg_n$ and this will obviously give us $f'$, so by induction we have shown that this may be done for any $n$.

\subsubsection{Question 27}
For every $b$ in the domain of $f$ there must be exactly one $a$ and $c$ such that $f(a) = b$ and $f(b) = c$. As the domain of $f$ is finite then there must be some $n\in\mathbb N$ such that $f^n(b) = b$. It follows then that if there is some $n$ such that $f^n(s) = t$ then there must also be some $k$ such that $f^k(t) = s$. By symmetry we also know that the converse is true. This means that either $O(s) = O(t)$ or the two are disjoint.

\subsubsection{Question 30}
Each orbit must be exactly of size $1$. This is because otherwise all $n$ such that $f^n = i$, would have to be a multiple of a number that is not $1$, and thus could not be any prime number.

\subsubsection{Question 32}
$g\in A(S)$ commutes with $f$ iff $g$ is closed on the set $\set{x_1,x_2}$.
\begin{proof}
	First we will show by cases that any $g$ that is closed on $\set{x_1,x_2}$ commutes with $f$, then we will show that no other set does so.

	\begin{itemize}
		\item Let $s,t \in \set{x_1,x_2}$ with $s\not=t$
		\begin{itemize}
			\item If $g(s) = s$, then $fg(s) = g(t) = t$ and $gf(s) = f(s) = t$.
			\item If $g(s) = t$, then $fg(s) = g(t) = s$ and $gf(s) = f(t) = s$.
		\end{itemize}
		\item Let $s \not\in\set{x_1,x_2}$, then $fg(s) = gf(s)$ as $f$ acts as the identity.
	\end{itemize}

	Now if $g$ is not closed on $\set{x_1,x_2}$ then lets say without loss of generality that $g(x_1) = s \not\in\set{x_1,x_2}$ it follows that $fg(x_1) = g(x_2)$ and $gf(x_1) = f(s) = s$. Now $g(x_2) \not= s$ as otherwise both $x_1$ and $x_2$ would map to the same element which is not possible.
\end{proof}

\subsection{Section 5}
\subsubsection{Question 1}
\subsubsection{Question 4}
\subsubsection{Question 7}
\subsubsection{Question 13}
\subsubsection{Question 17}

\end{document}