### 2018-04-06 Seminar/PushDownAutomata.tex
hi

\documentclass{article}
\input{../definitions}
\title{Push Down Automata}

\begin{document}

	\maketitle
	\question{1.} Make a PDA for the language of $a^nb^n$ that is $n$ $a$'s followed by $n$ $b$'s for any $n$. So ${\lambda, ab, aabb, aaabbb,\ldots}$

	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{2.} Make a PDA for a valid mathematical expression with symbols $+$, $-$, $*$, $/$, and $0$ through $9$.

	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{3.} Make a PDA that accepts any word with correctly matching parentheses, on the alphabet, $\set{a,b,c,(,)}$

	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip
	\bigskip

	\question{4.} Build up a language that you know as a grammer.

\end{document}