### 2019-04-14 SchoolCode/Foundations/3.3.tex

\documentclass{article}

\input{definitions}
\title{Foundations of Mathematics \\ Section 3.3}

\begin{document}

\maketitle

\question{3.3.4}

\question{a)}

Here our partition $\mathscr P$ on $\mathbb R$ may be defined as the following
\begin{enumerate}
\item $\All{x\in\coint01}{x \in P_x \in \mathscr P}$
\item $\All{P \in \mathscr P}{x \in P \implies \left( x + 1 \in P \And x - 1 \in P \right)}$
\end{enumerate}

\question{c)}
Here our partition $\mathscr P$ on $\mathbb R$ may be defined as the following
\begin{enumerate}
\item $\All{x \in \ccint{-\frac\pi2}{\frac\pi2}}{x \in P_x \in \mathscr P}$
\item $\All{x \in \ooint{-\frac\pi2}{\frac\pi2}}{\pi-x \in P_x}$
\item $\All{P \in \mathscr P}{x \in P \implies \left( x + 2\pi \in P \And x - 2\pi \in P \right)}$
\end{enumerate}

\question{d)}

Here our partition $\mathscr P$ on $\mathbb R$ may be defined as the following
\[
\All{x\in\coint0\infty}{\set{-x,x}\in\mathscr P}
\]

\question{e)}

Here our partition $\mathscr P$ on $\mathbb R^2$ may be defined as the following
\begin{enumerate}
\item $\mathscr P = \set{P_-, P_0, P_+}$
\item $P_- = \setbuilder{\pair xy \in \mathbb R^2}{xy < 0}$
\item $P_0 = \setbuilder{\pair xy \in \mathbb R^2}{xy = 0}$
\item $P_+ = \setbuilder{\pair xy \in \mathbb R^2}{xy > 0}$
\end{enumerate}

\question{3.2.5}

Here we define our partition $\mathscr P$ on $\set{1,i,-1,-i}$ as
\[\mathscr P = \set{\set{1,-1},\set{i,-i}}\]

\question{3.2.6}

Here we define our partition $\mathscr P$ on $S^2$ where $S = \set{1,i,-1,-i}$ as
\[\mathscr P = \setbuilder{\setbuilder{\pair xy\in S^2}{xy = s}}{s\in S}\]
This may more clearly be defined as the following, the two are equivalent.
\[\All{s\in S}{\setbuilder{\pair xy\in S^2}{xy = s} \in \mathscr P}\]

\end{document}